Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"

### 2018-06-21 results.tex
Updated paper with equivalence test results

Given the diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible. At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer for participants of all career stages. We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12). Only for GHW do we find a very weak suggestion that early-career researchers might agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)).

One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded. Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow a wide range of participants to participate fully.
Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$). Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$). Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ with respect to gender or racial/ethnic identity. For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$). These initial results indicate that future work on how demographics interact with hack week attendance may be fruitful.

>>>>

Given the diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible. At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer for participants of all career stages. We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12). Only for GHW do we find a very weak suggestion that early-career researchers might agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)). However, the confidence intervals on the measured effect sizes are wide, and thus testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size indicates that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).

One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded. Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow a wide range of participants to participate fully.
Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$). Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$). Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ with respect to gender or racial/ethnic identity. For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$). Equivalence tests for the hypothesis of a small or non-existent effect reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW). These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an initial indication that our careful facilitation strategies may be effective in fostering participation, and that future work on how demographics interact with hack week attendance may be fruitful.

### 2018-06-21 supplementary_material.tex
Updated paper with equivalence test results

We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size. For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size. We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V. For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound. If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.

@@

We found no $p$-values below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials} for any of the combinations of questions and demographic variables. In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication where future research may be fruitful. The results also discuss the present study's limitations.

@@

\begin{tabularx}{18cm}{p{3.5cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}

@@

\textbf{Question} & \multicolumn{7}{c}{\textbf{Astro Hack Week}} & \multicolumn{7}{c}{\textbf{GeoHackWeek}} & \multicolumn{7}{c}{\textbf{NeuroHackWeek}} \\ \midrule
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$  \\ \midrule

@@

New topics, tools, methods (a) & 	61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 		36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 		41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
Improvement of research (b) & 		61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 	35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 	40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
Better science (c) & 				59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 		35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 	40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\
Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 	40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule

@@

Improvement of research (b) & 		60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 		35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 	40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
Teaching (d) & 					61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 	36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 	40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
Connection building (e) & 			61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 		36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 	41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
Valued contributions (f) & 			57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 	34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 	38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule

@@

Improvement of research (b)  & 	59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 		34 &2 & 0.01 & 0.99 & 0.97 & 0 &		39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
Teaching (d) & 					61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 	34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 		39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
Connection building (e) & 			61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 		34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 	40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
Valued contributions (f) & 			58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 	32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 		37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\

>>>>

We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size. For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{}. We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V. For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound. If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.

@@

We found no $p$-values below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials} for any of the combinations of questions and demographic variables. In the main text, we place these null results in context, and refer to the tests that were near significance, coupled with a moderate to large effect size, as an indication where future research may be fruitful. The results also discuss the present study's limitations.

@@

\begin{tabularx}{24cm}{p{3.2cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}

@@

\textbf{Question} & \multicolumn{9}{c}{\textbf{Astro Hack Week}} & \multicolumn{9}{c}{\textbf{GeoHackWeek}} & \multicolumn{9}{c}{\textbf{NeuroHackWeek}} \\ \midrule
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ &$\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule

@@

New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0 & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0 & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0 & 0.32 & 0.17 \\
Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0 & 0.31 & 0.14 \\
Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0 & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0 & 0.28 & 0.12 \\
Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0 & 0.33 & 0.19 \\ \midrule

@@

Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0 & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0      & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0 & 0.30 & 0.05\\
Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0 & 0.30 & 0.05 & 		36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.37 & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0 & 0.36 & 0.11\\
Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0 & 0.28 & 0.03 & 		36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0      & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0 & 0.32 & 0.07 \\
Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0 & 0.28 & 0.03 &		34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0      & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0 & 0.35 & 0.10\\ \midrule

@@

Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0 & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0 & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0 & 0.32 & 0.06 \\
Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0 & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0 & 0.38 & 0.12 \\
Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0 & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0 & 0.33 & 0.07 \\
Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0 & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0 & 0.36 & 0.11\\
